Essays in \\Computational Management Science

### 2018-02-20 Chapters/Chapter01.tex
Several changes

Bitcoin (BTC) is the first digital currency, also known as digital money, internet money and cryptocurrency. It is the first currency based on cryptography techniques, distributed and decentralized, and with no central bank. Bitcoin is distributed since its ledger is public and is stored in thousands of computers. It is decentralized because there is no authority (or government) who decides its future --- any decision must be accepted by its community. The security of Bitcoin relies on digital signature technology and network agreement. While digital signature ensures ownership, i.e., the funds may only be spend by their owners, and nobody else; the network agreement both prevents double spending and ensures that all processed transactions have sufficient funds. In short, every transaction must spend only unspent funds, must have enough funds available, and must be signed by its owners, authorizing its execution. When all these requirements are met, the funds are transfered.

@@

---

In order to obtain $\mathbf{E}(T)$, we will use the formula $\frac{d}{ds}T^*(s)|_{s=0} = -\mathbf{E}(T)$, where $T^*$ is the Laplace transform of the pdf of T.

Let $f_T(x)$ be the pdf of T and $F_T(x)$ be the cdf of T, i.e., $F_T(x) = \int_0^x f_T(t) dt$. Thus, $\mathcal{L}\{F_T(x)\} = \frac{T^*(s)}{s}$.

\begin{align*}
	\mathcal{L}\{F_T(x)\} &= \int_0^\infty e^{-sx} F_T(x) dx \\
		&= \int_0^\infty e^{-sx} \left( 1 - e^{-\frac{x + a^2x^2/2}{\eta}} \right) dx \\
		&= \int_0^\infty e^{-sx} dx - \int_0^\infty e^{-sx} e^{-\frac{x + a^2x^2/2}{\eta}} dx \\
		&= \frac{1}{s} - \int_0^\infty e^{-sx -\frac{x + a^2x^2/2}{\eta}} dx \\
\end{align*}

Let $b=\frac{a}{\sqrt(2 \eta)}$ and $c = \frac{\sqrt(2 \eta)}{2a} \left( s + \frac{1}{\eta} \right)$, then $sx + \frac{x+a^2x^2/2}{\eta} = (bx + c)^2 - c^2$.
Let $y = bx+c \Rightarrow dy = bdx$.

\begin{align*}
	\int_0^\infty e^{-sx -\frac{x + a^2x^2/2}{\eta}} dx &= \int_0^\infty e^{-(bx+c)^2 + c^2} dx \\
		&= \int_c^\infty e^{-y^2 + c^2} \frac{dy}{b} \\
		&= \frac{e^{c^2}}{b} \int_c^\infty e^{-y^2} dy \\
		&= \frac{\sqrt{\pi}e^{c^2}}{2b} \erfc(c)
\end{align*}

$$ \mathcal{L}\{F_T(x)\} = \frac{1}{s} - \frac{\sqrt{\pi}e^{c^2}}{2b} \erfc(c) $$

Hence,
$$ T^*(s) = s \mathcal{L}\{F_T(x)\} = 1 - \frac{s\sqrt{\pi}e^{c^2}}{2b} \erfc(c) $$

As $\frac{d}{ds}c = \frac{\sqrt{2 \eta}}{2a}$,

\begin{align*}
\frac{d}{ds} T^*(s) &= - \frac{\sqrt{\pi}e^{c^2}}{2b} \erfc(c) - \frac{\sqrt{2 \eta \pi} sc e^{c^2}}{2ab} \erfc(c) + \frac{\sqrt{2 \eta \pi}s}{4ab} \\
\frac{d}{ds} T^*(s)|_{s=0} &= - \frac{\sqrt{2 \eta \pi}}{2a} e^{\frac{1}{2 \eta a^2}} \erfc\left(\frac{1}{a \sqrt{2 \eta}}\right) \\
\mathbf{E}[T] &= \frac{\sqrt{2 \eta \pi}}{2a} e^{\frac{1}{2 \eta a^2}} \erfc\left(\frac{1}{a \sqrt{2 \eta}}\right) \\
\end{align*}

$\mathbf{E}[T]$ is ploted for $a \in (0, 0.1]$ at Figure \ref{fig-bitcoin-H-increase}. The plot interval is really small because $a=1$ would imply doubling the hash rate after 1 second.

\textcolor{red}{I do not know how to interpret this.}

@@

There are many possible ways to attack the Bitcoin (REF). In this section, we are interested in a particular attack: the submarine attack.

In the submarine attack, the attacker's goal is to spend some Bitcoins and, after a while, get their Bitcoins back. Thus, the attacker sends some funds to the victim, they wait 6 confirmations, and, finally, the victim confirms the funds. But, then, the attacker get their funds back propagating new blocks which accept a double spending transaction. In order to work, the attacker must propagate more blocks than the network has mined in the same period, becoming the longest chain. Hence, we would like to understand what the odds are that the attacker will be successful.

In order to maximize their odds, the attacker must start to mine the new blocks as soon as they send the funds to the victim. In this moment, it starts to mine in the head of the blockchain, just like the rest of the network. So, the attacker and the network are in exactly the same point.

@@

>>>>

Bitcoin (BTC) is the first digital currency, also known as digital money, internet money, and cryptocurrency. It is the first currency based on cryptography techniques, distributed and decentralized, and with no central bank. Bitcoin is distributed since its ledger is public and is stored in thousands of computers. It is decentralized because there is no authority (or government) who decides its future --- any decision must be accepted by its community. The security of Bitcoin relies on digital signature technology and network agreement. While digital signature ensures ownership, i.e., the funds may only be spend by their owners, and nobody else; the network agreement both prevents double spending and ensures that all processed transactions have sufficient funds. In short, every transaction must spend only unspent funds, must have enough funds available, and must be signed by its owners, authorizing its execution. When all these requirements are met, the funds are transfered.

@@

%In order to obtain $\mathbf{E}(T)$, we will use the formula $\frac{d}{ds}T^*(s)|_{s=0} = -\mathbf{E}(T)$, where $T^*$ is the Laplace transform of the pdf of T.
%
%Let $f_T(x)$ be the pdf of T and $F_T(x)$ be the cdf of T, i.e., $F_T(x) = \int_0^x f_T(t) dt$. Thus, $\mathcal{L}\{F_T(x)\} = \frac{T^*(s)}{s}$.
%
%\begin{align*}
%	\mathcal{L}\{F_T(x)\} &= \int_0^\infty e^{-sx} F_T(x) dx \\
%		&= \int_0^\infty e^{-sx} \left( 1 - e^{-\frac{x + a^2x^2/2}{\tau}} \right) dx \\
%		&= \int_0^\infty e^{-sx} dx - \int_0^\infty e^{-sx} e^{-\frac{x + a^2x^2/2}{\tau}} dx \\
%		&= \frac{1}{s} - \int_0^\infty e^{-sx -\frac{x + a^2x^2/2}{\tau}} dx \\
%\end{align*}
%
%Let $b=\frac{a}{\sqrt(2 \tau)}$ and $c = \frac{\sqrt(2 \tau)}{2a} \left( s + \frac{1}{\tau} \right)$, then $sx + \frac{x+a^2x^2/2}{\tau} = (bx + c)^2 - c^2$.
%Let $y = bx+c \Rightarrow dy = bdx$.
%
%\begin{align*}
%	\int_0^\infty e^{-sx -\frac{x + a^2x^2/2}{\tau}} dx &= \int_0^\infty e^{-(bx+c)^2 + c^2} dx \\
%		&= \int_c^\infty e^{-y^2 + c^2} \frac{dy}{b} \\
%		&= \frac{e^{c^2}}{b} \int_c^\infty e^{-y^2} dy \\
%		&= \frac{\sqrt{\pi}e^{c^2}}{2b} \erfc(c)
%\end{align*}
%
%$$ \mathcal{L}\{F_T(x)\} = \frac{1}{s} - \frac{\sqrt{\pi}e^{c^2}}{2b} \erfc(c) $$
%
%Hence,
%$$ T^*(s) = s \mathcal{L}\{F_T(x)\} = 1 - \frac{s\sqrt{\pi}e^{c^2}}{2b} \erfc(c) $$
%
%As $\frac{d}{ds}c = \frac{\sqrt{2 \tau}}{2a}$,
%
%\begin{align*}
%\frac{d}{ds} T^*(s) &= - \frac{\sqrt{\pi}e^{c^2}}{2b} \erfc(c) - \frac{\sqrt{2 \tau \pi} sc e^{c^2}}{2ab} \erfc(c) + \frac{\sqrt{2 \tau \pi}s}{4ab} \\
%\frac{d}{ds} T^*(s)|_{s=0} &= - \frac{\sqrt{2 \tau \pi}}{2a} e^{\frac{1}{2 \tau a^2}} \erfc\left(\frac{1}{a \sqrt{2 \tau}}\right) \\
%\mathbf{E}[T] &= \frac{\sqrt{2 \tau \pi}}{2a} e^{\frac{1}{2 \tau a^2}} \erfc\left(\frac{1}{a \sqrt{2 \tau}}\right) \\
%\end{align*}
%
%$\mathbf{E}[T]$ is ploted for $a \in (0, 0.1]$ at Figure \ref{fig-bitcoin-H-increase}. The plot interval is really small because $a=1$ would imply doubling the hash rate after 1 second.
%
%\textcolor{red}{I do not know how to interpret this.}

@@

There are many possible ways to attack the Bitcoin (REF). In this section, we are interested in a particular attack: the double spending attack.

In the double spending attack, the attacker's send some funds to the victim, let's say a merchant. They wait for $k$ confirmations of the transaction, and the victim delivers the good or the service to the attacker. Then, the attacker mine enough blocks with a conflicting transaction, double spending the funds which was sent to the victim. If the attacker is successful, the original transaction will be \textit{erased} and the victim will be left with no funds at all. In order to be successful, the attacker must propagate more blocks than the network in the same period, propagating a chain longer than the main chain. Hence, we would like to understand what the odds are that the attacker will be successful. This attack was originally discussed by \citet{nakamoto2008bitcoin}.

In order to maximize their odds, the attacker must start to mine the new blocks as soon as they send the funds to the victim. In this moment, it starts to mine in the head of the blockchain, just like the rest of the network. So, in the beginning, the attacker and the network are in exactly the same point.

@@

When a disagreement between miners' rules happens, that is referred as either a hard-fork or a soft-fork. It is said to be a soft-fork when the rules are backward compatible, and a hard-fork when the rules are not backward compatible. In general, a hard-fork relaxes the constraints, while the latter hardens them.

Suppose that the miners' are split in two groups, $G_1$ and $G_2$, with different rules. Let's say $H_1$ and $H_2$ are their hash rate, respectively. We have two different scenarios to analyze: (i) when neither of them accepts other's blocks; and (ii) when $G_2$ accepts $G_1$'s blocks, but not the other way around.

Scenario (i) is easy to analyze, because the network would just split and, after a while, both difficulties will be adjusted. Then, they will be just like two different Bitcoin networks.

Scenario (ii) is more trick. As $G_2$ accepts $G_1$ blocks, their hashrate will matter. If $H_1 > H_2$, then $G_2$ will frequently skip their blockchain, because $G_1$'s blockchain will be longer most of the time. If $H_1 < H_2$, then $G_2$ may have its own blockchain after a while --- a true fork. But what would happen if $H_1$ keeps increasing and eventually gets larger than $H_2$? It if happens for sufficient time, the whole $G_2$'s blockchain may be discarded in order to move to $G_1$'s blockchain when $G_1$'s gets longer.

\begin{theorem}
	When $H_1 > H_2$, the 
\end{theorem}

%What would happen if, after a while, the groups agree again. How would the blockchain merge?

### 2018-02-20 Chapters/Chapter02.tex
Fix sdm2.tex path
Update on Overleaf.
Several changes

\input{sdm2.tex}

@@

@@

The framework implements three main scanner algorithms: linear scanner, thread scanner, and OpenCL scanner. The linear scanner runs in a single core and is the slowest one. It was developed only for testing purposes. The thread scanner runs at the CPU in multiple threads (and our recommendation is to use the number of threads equals to twice the number of cores). The OpenCL scanner runs at the GPUs and support multiple devices. The speed of a scan depends on the processors and GPU devices, thus the best approach to choose which scanner is best for one's setup is to run a benchmark test.

The hamming distance between two bitstrings is calculated counting the number of ones in the XOR of the bitstrings, i.e., $d(x, y) = \text{number of ones in} x \oplus y$. In order to count the number of ones of the XOR, the framework uses a pre-calculated table of distances --- by default, it is used a 16-bit table which uses 65MB of RAM, but one may increase to a 32-bit table which uses 4GB of RAM. Finally, to calculate the distance between to bitstrings, it sums the distance of each 16-bit part of the bitstrings, i.e., $d(x[0:N-1], y[0:N-1]) = d(x[0:15], y[0:15]) + d(x[16:31], y[16:31]) + \dots$. As each distance is calculated in O(1), this algorithm runs in O($\lceil bits/16 \rceil$).

Internally, the address space is stored in an array of addresses, which, in turn, is stored in an array of 64-bit integers (Fig. \ref{tab:hl-addresses-detail}). As the addresses and their 64-bit integers are contiguous in RAM or file, there is a performance gain since they are loaded in chunks.

@@

\section{Proposal:  Framework Validation}

In order to check whether the framework works as expected and according to the model proposed by \citep{Kanerva1988}, we run simulations which were compared to the theoretical analysis conducted some decades ago.

@@

\section{Proposal: Performance Results}

@@

\chapter{Conclusion}
Sparse Distributed Memory is a viable model of human memory, yet it does require researchers to (re-)implement a number of parallel algorithms in different architectures.

@@

>>>>

\input{./Chapters/sdm2.tex}

@@

\section{Read and write operations}

@@

Counters may have overflow protection depending on compiling options. By default, there is no overflow check in the counters, since each counter is a signed 32-bit integer --- which can store any number integer -2,147,483,648 and 2,147,483,647 --- and they will not overflow with less than $2^31-1$ divided by the number of activated hard-locations. When $N=1,000$, $H=1,000,000$, and $r=451$, the average number of activated hard-locations is 1,000 and it would require at least one million writes to overflow a counter.

\section{Implementation details}

\subsection{The distace between two bitstrings}

The hamming distance between two bitstrings is calculated counting the number of ones in the XOR of the bitstrings, i.e., $d(x, y) = \text{number of ones in} x \oplus y$. In order to count the number of ones of the XOR, the framework uses a pre-calculated table of distances --- by default, it is used a 16-bit table which uses 65MB of RAM, but one may increase to a 32-bit table which uses 4GB of RAM. Finally, to calculate the distance between two bitstrings, it sums the distance of each 16-bit part of the bitstrings, i.e., $d(x[0:N-1], y[0:N-1]) = d(x[0:15], y[0:15]) + d(x[16:31], y[16:31]) + \dots$. As each distance is calculated in O(1), this algorithm runs in O($\lceil bits/16 \rceil$).

The address space is stored in an array of addresses, which, in turn, is stored in arrays of 64-bit integers (Fig. \ref{tab:hl-addresses-detail}). As the addresses and their 64-bit integers are contiguous in RAM or file, there is a performance gain because processors load them in chunks.

@@

\subsection{Scanning for activated hard-locations}

Scanning for the activated hard-locations is a problem similar to well-known problems in computation geometry called ``range reporting in higher dimensions''. In this case, none of the known algorithms is able to solve our problem faster than $O(n)$. The algorithm which seems to best fit in our problem consumes $O(n)$ space and runs in $O(\log^d(n))$ \citep{chazelle1988functional}, which is really slower than $O(n)$ when, for instance, $n=1,000,000$ and $d=1,000$. For a review of the algorithms, see \citet{chan2011orthogonal}.

In 2014, there was published a solution to fast search in hamming space which seems applicable to our problem \citet{norouzi2014fast}.

It is intriguing that none of those algorithms is able to solve our scanning problem. The idea behind those algorithms is roughly to split the search space in half each step, which would take $O(\log(n))$ to go through the whole space. But this approach does not work in our case because of the high number of dimensions (i.e., 1,000) and because the hardlocations' addresses are randomly sampled from the ${0, 1}^d$ space. So, each addresses' bit itself splits the hardlocations in half, but it does not split the search space in half since both halves still must be convered. For instance, let's say we have 1,000 dimensions with 1,000,000 hardlocations, and we are scanning within a circle with radius 451, then after checking the first bit we have two cases: (i) for the half with the same first bit, we must keep scanning with radius 451; and (ii) for the half with a different first bit, we must keep scanning with radius 450. The search space has not been split in half because both halves have been covered (and one of them should have been skipped).

Hence, as our best approach approach seems to be scanning through all hardlocations, we may distribute the scan into many tasks which will be executed independently. The tasks may be executed in different processes, threads, or even computers. They may also run in the CPU or in the GPU. In this case, we may take into account both the time required to distribute the tasks and the time to receive their results.

The framework implements three main scanner algorithms: linear scanner, thread scanner, and OpenCL scanner. The linear scanner runs in a single core, is the slowest one, and was developed only for testing purposes; the thread scanner runs at the CPU in multiple threads sharing memory (and our recommendation is to use the number of threads equals to twice the number of CPU cores); and the OpenCL scanner runs in multiple GPU cores and support multiple devices. The speed of a scan depends on the CPU and GPU devices, thus the best approach to choose which scanner is best for one's setup is to run a benchmark.

The OpenCL must be initialized, which just copies the hardlocations' addresses to the GPU's memory. Then, many scans may be executed with no necessity to upload the addresses again.

\section{Framework Validation}

The framework has been validated comparing its results with the expected results from \citet{Kanerva1988}. Thus, we run simulations which were then compared to the theoretical analysis conducted some decades ago.

SHOW THE RESULTS --- Kanerva's tables/figures vs Simulation.

@@

\section{Applications}

\subsection{Noise filter}

We developed a noise filter using a SDM.

\subsection{Classification filter}

\section{Performance Results}

@@

\chapter{Conclusion}

Sparse Distributed Memory is a viable model of human memory, yet it does require researchers to (re-)implement a number of parallel algorithms in different architectures.

@@

\chapter{Appendix}
\section{Generating Kanerva's table 7.3}

\includepdf[
  pages=-,
  pagecommand={\pagestyle{headings}},
  %addtotoc={1,section,1,Quilling Shapes,sec:shapes}
]{./Chapter02-SDM/Kanerva-Table-7.3/Kanerva-Table-7_3.pdf}

### 2018-02-20 Chapters/Conclusion.tex
Fix compiling error
Update on Overleaf.
Several changes

\noindent If anything good can ever be said about the second world war, it might be this: the war effort sparked a massive number of scientific fields.

>>>>

\begin{flushright}{\slshape
	Software is eating the World.} \\ \medskip
    --- Mark Andreesen \citep{andreesen:2011}
\end{flushright}
\bigskip
\bigskip

\noindent Modern management and high technology interact in multiple, profound, ways.  Software, in particular, seems to have an immense power of entering arenas which seemed, at some point, to require either specific hardware or the skill of humans.  One of the members of this thesis committee, Dr. Nichols, will participate through teleconferencing over the open web, with no use of hardware specific for the task.  The corporate biography of Tonny Martins, President of IBM Brasil, mentions his successes with blockchain, AI and cognitive technology… as an executive, not as a research scientist or specialized engineer (https://www-03.ibm.com/press/br/pt/biography/53561.wss). Professor Andrew Ng tells students at Stanford’s Graduate School of Business that ``AI is the new electricity'' \citet{AndrewNgAI}, as his way to emphasize the potential transformational power of the technology.  It is not impossible that a purely digital form of money may exist.  It is not impossible that machines may become intelligent.  Moreover, it is not impossible that these two processes may have already begun.

It is worthwhile, in this concluding section, to reflect on some ideas on what this thesis is and what it is that we, as computational management scientists, can obtain from this sort of study.  Clemenceau once said that ``war is too important to be left to the generals''.  I believe it is not far-fetched to state that technology has become too important to be isolated to the realm of computer science, or engineering, or applied mathematics, or any single discipline.  The emergence of scientific journals with names such as Computational Management Science; INFORMS Journal on Computing; Ledger; Computational Statistics; ACM Transactions on Economics and Computation, and so forth, show that there are growing communities deeply interested in the intersection of business and the computer sciences.  To whom, for instance, does the OpenAI project belong?  To computing or to business?  Recall that the project was created as a risk-management strategy against the far-fetched, but not impossible, possibility of having machines yielding too much power.  What about corporations like Uber?  AirBnB?  Imagine a new method that increases profits by 50\% at a tech company.  Should this method, if implementable as an algorithm (like PageRank \citet{brin1998anatomy}) or a data structure (like a blockchain) be discussed in conferences of `computer science' or `business'? It seems quite arbitrary to name a single group, as a whole new ecosystem seems to be emerging within those two. That is why this thesis is computational and why it is business.  This work explores topics that seem, on the surface, to belong to computer science, but their applicability and impact to businesses seem too large to be delegated away, something ``for the nerds in the fifth floor''.  As technical decisions become central to the organization of man's life, the technician becomes the visionary, the innovator, the decision-making arbiter, sometimes the billionaire.

We have started this study with two possible forms of organization of a purely digital money system; a blockchain and a directed acyclic graph; we then moved to an analysis of the adoption of new technologies.

The possibility that there will be some form of purely digital money has become very real.  Consider, just as a matter of comparison, Brazil's most important company: Petrobrás.  As of this writing, the ``market cap'' of Ethereum exceeds that of Petrobras by ten billion dollars, while Bitcoin's is valued at more than double of Petrobras (195B usd vs 83B usd).  These technologies should, at a minimum, be taken seriously.

%P.5. CryptoPonzi #2. Dag/Tangle Results. Some speculative possibilities can be mentioned here:

We have explored Kanerva’s Sparse Distributed Memory.   In AI, SDM seems to be a particularly interesting area for study.  The model plausibly reflects a number of well known aspects of psychology and neuroscience.  For example, neurons can easily compute the address decoding scheme of the system.  Neurons are fragile and may be lost, whereas the information remains preserved, due to the distributed character of the model.  The ``tip-of-the-tongue'' behavior emerges naturally, and so does Miller’s magic number.

There are three contributions made on SDM:  First, I have illuminated a discrepancy between Kanerva’s theoretical model and the real system dynamics; Also, we have seen that pattern classification through supervised learning is possible without presuming any new SDM mechanism. This is in contrast with the literature, that presumes additional mechanisms, like genetic algorithms, to account for supervised learning.  Finally, we now have a tested open-source framework that offers parallelism and can become a de-facto standard in SDM research.  The framework (i) carefully reproduces crucial figures from Kanerva’s theoretical book; (ii) shows how noise filtering and (iii) supervised learning can be done, and, through the use of (iv) Jupyter Notebooks, enables the reader to easily reproduce all the results on their own machines. This respects all constraints posed by Robert M. French in his article on Computational Modeling in Cognitive Science: A Manifesto for Change \citep{TOPS:TOPS1206}.

The ability to rapidly reproduce results, and to build on prior work, is, I believe, fundamental to modern science.  Consider, for instance, the groundbreaking successes in the arena of deep learning.  Having standard computer libraries to work with has brought together a community, which reinforces the system, as users also gradually improve these libraries. It may be possible to achieve new results with multiple layers of a SDM, yet, having to start development from scratch takes a large opportunity cost from most scientists --- especially those who are less concentrated on the computer science aspects, but still would be able to contribute meaningfully.

Finally, we have studied how variations of the Bass Model may reflect systems or technologies that may wither in time. Though some innovations, such as the radio, have gained widespread use in a sustainable form... One may want to review the Bass model when one is concerned with rapidly-evolving technological ecosystems.  Hardly anyone remembers the names AskJeeves, World Wide Web Worm, Lycos, WebCrawler, or AltaVista, early web search engines; later replaced, in the market and by the market, by the almost unnoticed url http://google.stanford.edu \citep{brin1998anatomy}.

Another possibility would be to compare the proposed model with a computation of the momentum of Metcalfe’s law in between competitors.  As the reader may remember, Metcalfe’s law states that the value of a network grows $O(n^2)$ with $n$ being the number of network nodes.   If the proposed model and Metcalfe’s network effects reflect reality, then there could be an integrated mathematical model that explains and represents both Metcalfe’s law and the variation of the Bass model presented herein.

With this, I submit this thesis in the hope that all readers, present and future, may find the aforementioned studies as useful, genuine, and legitimate contributions to the thriving field of Computational Management Science.

### 2018-02-20 Chapters/Introduction.tex
Several changes

The possibility of having a completely digital financial system without the overheads of traditional banking systems has appeared with the release of Bitcoin and similar blockchain technologies.  This field questions numerous traditional assumptions in computer science, record-keeping, banking \& finance, and economic inclusion. The seminal work of \citet{nakamoto2008bitcoin} described the architecture of Bitcoin, a peer-to-peer electronic cash system, also known as cryptocurrency. Bitcoin's currency ledger is public and stored in a blockchain across thousands of computers. Even so, no one is able to spend either somebody else's funds nor to double spend their own funds. In order to be confirmed, each transaction must be both digitally signed by the owner of the money and the funds verified in the blockchain by Bitcoin's miners.

The question of whether Bitcoin (or related works) can scale to billions of people is, however, far from settled.

One of the interesting parts of Bitcoin are the incentives. On one hand, users have incentive to use Bitcoin because the fees are small, the money transfer is quick, and the currency issuance rate is well known. On the other hand, the miners have incentive to be part of Bitcoin's network because new coins are found every ten minutes.

The impact of Bitcoin in society -- and hence in the companies and the government -- has been increasing every day. People are increasingly using Bitcoin to exchange money and transfer money overseas. Companies are looking into Bitcoin as an alternative to reduce banking fees. The poor may be included in the finance system through Bitcoin. People may hedge their assets against their governments' money issuance and inflation---as in the case of Venezuela.

@@

Note that these problems have been a deliberate decision from the current developers of the ``bitcoin-core'' developers, which believe that is it risky to increase the the blocksize (in which all transactions are stored).  It is not known whether a blocksize, say, of 1Gb, would be feasible to sustain the decentralization of the network.

@@

Artificial intelligence has been disrupting many businesses. Uber is able to handle hundred of thousands of requests. Amazon optimizes the location of each product based on demand. Netflix increases the quality of their services offering movies specific to the taste of each customer. Spotify learns which kind of music users like the most and suggests playlists and radios. Banks prevent fraud classifying which patterns are strange to their customers.

>>>>

The possibility of having a completely digital financial system without the overheads of traditional banking systems has appeared with the release of Bitcoin and similar blockchain technologies.  This field questions numerous traditional assumptions in computer science, record-keeping, banking \& finance, and economic inclusion. The seminal work of \citet{nakamoto2008bitcoin} described the architecture of Bitcoin, a peer-to-peer electronic cash system, also known as cryptocurrency. Bitcoin's currency ledger is public and stored in a blockchain across thousands of computers. Even so, no one is able to spend either somebody else's funds nor to double spend their own funds. In order to be confirmed, each transaction must be both digitally signed by the owner of the money and the funds verified in the blockchain by Bitcoin's miners. The question of whether Bitcoin (or related works) can scale to billions of people is, however, far from settled.

One of the interesting parts of Bitcoin are the incentives. On one hand, users have incentive to use Bitcoin because the fees are small, the money transfer is quick and global, and the currency issuance rate is well known. On the other hand, the miners have incentive to be part of Bitcoin's network because new coins are found every ten minutes. These incentives put the community together and has maintened Bitcoin alive.

The impact of Bitcoin in society -- and hence in the companies and the government -- has been growing every day. People are increasingly using Bitcoin to exchange money and transfer money overseas. Companies are looking into Bitcoin as an alternative to reduce banking fees. The poor may be included in the finance system through Bitcoin. People may hedge their assets against their governments' money issuance and inflation --- as in the case of Venezuela.

@@

Note that these problems have been a deliberate decision from the current developers of the ``bitcoin-core'', which believe that is it risky to increase the the blocksize (in which all transactions are stored).  It is not known whether a blocksize, say, of 1GB, would be feasible to sustain the decentralization of the network.

@@

Artificial intelligence has been disrupting many businesses. Uber is able to handle hundred of thousands of requests. Amazon optimizes the location of each product based on demand. Netflix increases the quality of their services offering movies specific to the taste of each customer. Spotify learns which kind of music users like the most and suggests playlists. Banks prevent fraud classifying which patterns are strange to their customers.

### 2018-02-20 Chapters/sdm2.tex
Add missing files
Fix typo
Several changes

\begin{figure}[h]
\centering\includegraphics[scale=0.75]{images/p_circle_r.pdf}

@@

\begin{figure}[h]
\centering\includegraphics[scale=0.75]{images/p1_inter_p2.pdf}

@@

Kanerva also suggests the selection of a radius that will activate, on average, one one thousandth of the sample, which is 1,000 hard-locations for a sample of one million addresses. In order to achieve his suggestion, a 1,000-dimension memory uses an access radius $r=451$, and a 256-dimensional memory, $r=103$.

\begin{figure}[h]
\centering\includegraphics[scale=0.75]{images/hardlocations.pdf}

@@

\begin{figure}[h]
\centering\includegraphics[scale=0.75]{images/p1_p2_iter_read.pdf}

@@

\begin{figure}[h]
\centering\includegraphics[scale=0.75]{images/p1_after_write.pdf}

@@

SDM memory or an extension of it has been used in a lot of applications, helping to solve specific problems. \citet{Snaider2011} extended SDM to efficiently store sequences of vectors and trees.  \citet{Rajesh1998} used SDM in an autonomous robot. \citet{Meng2009} modified SDM to clean patterns from noisy inputs.

@@

>>>>

\begin{figure}[h]
\centering\includegraphics[scale=0.75]{./images02/p_circle_r.pdf}

@@

\begin{figure}[h]
\centering\includegraphics[scale=0.75]{./images02/p1_inter_p2.pdf}

@@

Kanerva also suggests the selection of a radius that will activate, on average, one one thousandth of the sample, which is 1,000 hard-locations for a sample of one million addresses. In order to achieve his suggestion, a 1,000-dimension memory uses an access radius $r=451$, and a 256-dimensional memory, $r=103$. We think that a 256-dimensional memory may be important because it presents conformity to Miller's magic number \citep{Linhares2011}.

\begin{figure}[h]
\centering\includegraphics[scale=0.75]{./images02/hardlocations.pdf}

@@

\begin{figure}[h]
\centering\includegraphics[scale=0.75]{./images02/p1_p2_iter_read.pdf}

@@

\begin{figure}[h]
\centering\includegraphics[scale=0.75]{./images02/p1_after_write.pdf}

@@

SDM memory or an extension of it has been used in many applications, helping to solve specific problems. \citet{Snaider2011} extended SDM to efficiently store sequences of vectors and trees.  \citet{Rajesh1998} used SDM in an autonomous robot. \citet{Meng2009} modified SDM to clean patterns from noisy inputs.

@@

A member of my Master's committee, professor Paulo Murilo, has proposed a generalized reading operation (personal communication), which covers both Kanerva's and Chada's read --- and opens a new venue of potential discoveries. He proposed summing all hard-location counters raised to the power of $z$ while holding the original sign of the counter (positive or negative). Thus, Kanerva's read would be the same as $z=1$, while Chada's would be the same as $z=0$. Hence, we might explore how SDM would behave with other values of $z$, such as 0.5, 2, and 3.

### 2018-02-20 FrontBackmatter/Acknowledgments.tex
Several changes

I must, first and foremost, give thanks for the complete and unwavering support and love from all my family. My beloved wife Patrícia, my wonderful parents Reynaldo and Angelina, my lovely sister Flávia, my sweet grandmas Geny and Rosaria, and all others who will always be remembered in my heart. Thank you for your patience and for understanding that I was physically far away but very close in heart. A special thanks to my dad, Reynaldo, who always taught me about life and who first awakened my passion for technology and engineering.

Thanks to my wife, Patrícia, for sharing the dreams, the passion, and the craziness, for all her caring and love; and for discussing about this work over and over again. Especially, for understanding my absence.

Thanks to my awesome advisor, Alexandre Linhares, who became a dear friend during this journey.

Thanks to my dear friends Daniel Chada, Layla Mendes, Renato Kogeyama, Luiz Sacramento, Kaillen Givigi, Jamil Chivitarese, Lucia Barros, Daniel Modenesi, André Luiz, Bernardo Machado, always available for serious talking, laughing, and sharing their wide and deep knowledge, even at lunchtime and weekends.

Thanks to Celene.

Thanks to Claudio Abreu.

Thanks to professor Luiz Antônio Joia, who trusted me in the first place. Additionally, thanks to EBAPE / FGV for their excellent teachers and infrastructure.

>>>>

%I must, first and foremost, give thanks for the complete and unwavering support and love from all my family. My beloved wife Patrícia, my wonderful parents Reynaldo and Angelina, my lovely sister Flávia, my sweet grandmas Geny and Rosaria, and all others who will always be remembered in my heart. Thank you for your patience and for understanding that I was physically far away but very close in heart. A special thanks to my dad, Reynaldo, who always taught me about life and who first awakened my passion for technology and engineering.

%Thanks to my wife, Patrícia, for sharing the dreams, the passion, and the craziness, for all her caring and love; and for discussing about this work over and over again. Especially, for understanding my absence.

%Thanks to my awesome advisor, Alexandre Linhares, who became a dear friend during this journey.

%Thanks to my dear friends Daniel Chada, Layla Mendes, Renato Kogeyama, Luiz Sacramento, Kaillen Givigi, Jamil Chivitarese, Lucia Barros, Daniel Modenesi, André Luiz, Bernardo Machado, always available for serious talking, laughing, and sharing their wide and deep knowledge, even at lunchtime and weekends.

%Thanks to Celene.

%Thanks to Claudio Abreu.

%Thanks to professor Luiz Antônio Joia, who trusted me in the first place. Additionally, thanks to EBAPE / FGV for their excellent teachers and infrastructure.
